A weakly structured stem for human origins in Africa

### 2022-07-15 paper.tex

Archaeological sites from the Middle Stone Age (approx. 300-40ka) are widely
distributed across Africa, and are particularly well represented in the
northern, eastern and southern parts of the continent. Similarly, fossils 
such as those from the sites of Jebel Irhoud, Morocco \citep{Hublin2017-cq}, Herto, Ethiopia
\citep{White2003-bk} and Klasies River, South Africa \citep{Deacon1995-rx} demonstrate that 
derived \emph{Homo sapiens} anatomical features were also present across the
continent during this period. It has been difficult to reconcile these lines of
evidence with evidence from genomics, which have suggested a predominantly
tree-like model of recent population divergence from a single ancestral
population. It is unclear whether fossil specimens and archaeological sites
represent populations which contributed to our ancestors as population
precedents, or were local ``dead-ends'' from which contemporary \emph{Homo
sapiens} do not descend. 
Recently,
attempts to reconcile genetic and
paleoanthropological data include proposals for a Pan-African origin of
\emph{Homo sapiens} by which populations in many regions of the continent
contributed to the formation of \emph{Homo sapiens} beginning at least 300ka
\citep{Stringer2016-mj,Scerri2018-nl,Scerri2019-xg}.

>>>>

Decades of study of human genome variation have suggested a predominantly tree-like model of recent population divergence from a single ancestral population in Africa. It has been difficult to reconcile this finding with the fossil and archaeological record of human occupation across the vast African continent. For example, fossils such as those from the sites of Jebel Irhoud, Morocco \citep{Hublin2017-cq}, Herto, Ethiopia \citep{White2003-bk} and Klasies River, South Africa \citep{Deacon1995-rx} demonstrate that derived \emph{Homo sapiens} anatomical features were present across the continent 300-100ka. Archaeological sites from the Middle Stone Age, which has been associated with \emph{Homo sapiens} at some sites, are also widely distributed across Africa. It is unclear whether fossil specimens and archaeological sites represent populations which contributed to our ancestors as population precedents, or were local ``dead-ends'' from which contemporary \emph{Homo sapiens} do not descend. Recently, attempts to reconcile genetic and paleoanthropological data include proposals for a Pan-African origin of \emph{Homo sapiens} by which populations in many regions of the continent contributed to the formation of \emph{Homo sapiens} beginning at least 300ka \citep{Stringer2016-mj,Scerri2018-nl,Scerri2019-xg}.
